---
title: "Song Data Project"
author: "Kristian Abad, Steven Truong, Nicole Magallanes"
date: "2/13/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(tidyr)
library(tidyverse)
```
# Preprocessing
```{r}
data = read_csv("song_data.csv")
head(data)
```
One challenge we need to figure out is addressing the following cases in our data, if there are any:\

* Remixes
* Remasters
* Single Versions
* Same name but different artists?
\
I think maybe we can leave remixes possibly treating them as reimaginings of songs or somewhat to the same vein that songs have samples from other tracks are in themselves a separate track. Maybe the more difficult is dealing with the other cases. An example that comes to mind is "Smooth Operator" by Sade (seems like only one of the 3 versions is in the data). There's a single version, a remastered version, and I believe an album version where there's an immediate difference between the remastered and album version.
\
I think the duplicated() function finds exact duplicates of rows.
```{r}
duplicates <- data[duplicated(data),]
duplicates
```
Just testing some cases here...while scrolling through on kaggle, I just picked a random duplicate song to test
```{r}

duplicates %>%
  filter(song_name == 'Zombie')
```
Here's an interesting case where we have 2 of the same rows and 1 with a remix with a track called "8 Letters"
```{r}
duplicates %>%
  filter(song_name == '8 Letters')

duplicates %>%
  filter(song_name == '8 Letters - R3HAB Remix')
```
So it looks like it just picks up exact duplicates and we'll need to figure out what we're going to do with other cases.


